---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setoptions,echo=TRUE}

```

## Loading and preprocessing the data
Load libraries and delete variables

```{r}
library(plyr)
library(dplyr)
library(data.table)
library(chron)
library(lattice)
# library(lubridate)
rm(list=ls())
```
Create data directory (if it does not exist previously) and unzip the .zip data
file in the "data" directory.
```{r}
# Create data folder if it does not exist
if (!file.exists("data")) {
        dir.create("data")
}
# Unzip file in "data" directory
unzip(zipfile = "activity.zip", exdir = "./data")
```

The data frame is created with fread and the data structure and type of 
each variable is checked.
```{r}
# Create a "Data Frame" with "fread"
df <- fread("./data/activity.csv")
head(df)
str(df)
summary(df)
```

## What is mean total number of steps taken per day?
### *Calculate the total number of steps taken per day*
```{r}
# For this part of the assignment, you can ignore the missing values in the dataset.
# Note the use of the '.' function to allow date be used without quoting
df_day <- ddply(df, .(date), summarize, 
                sum = sum(steps, na.rm = TRUE), 
                mean = mean(steps, na.rm = TRUE),
                median = median(steps, na.rm = TRUE)
                )

df_day
```
### *Make a histogram of the total number of steps taken each day*
```{r histogram_01}
hist(df_day$sum,
     main="Histogram of total number of steps taken each day", 
     xlab = "total number of steps taken each day",
     col="#2166AC")

dev.copy(png, file = "figure/hist_steps_day.png")
dev.off()
```
### *Calculate and report the mean and median of the total number of steps taken per day*
```{r}
# Calculate and report the mean and median of the total number 
# of steps taken per day
mean_step_day <- mean(df$steps, na.rm = TRUE)
median_step_day <- median(df$steps, na.rm = TRUE)
```
The **mean** of the total number of steps taken per day 
is **`r mean_step_day`**.


The **median** of the total number of steps taken per day is
**`r median_step_day`**.

## What is the average daily activity pattern?
### Make plot of the 5-minute interval *vs* average number of steps taken
Calculation of average number of steps per 5-minute interval per day
```{r}
# Make a time series plot type = "l" of the 5-minute interval (x-axis) and
# the average number of steps taken, averaged across all days (y-axis)

df_minute <- ddply(df, .(interval), summarize, 
                mean = mean(steps, na.rm = TRUE),
                )

head(df_minute)
```
Time series plot type = "l" of the 5-minute interval (x-axis) and 
the average number of steps taken, averaged across all days (y-axis)

```{r plot_line_mean_interval}
plot(x=df_minute$interval,
     y=df_minute$mean,
     type="l",
     main="Average number of steps per interval", 
     xlab = "5-minute interval",
     ylab = "Average number of steps",
     col="#2166AC",
     frame.plot = FALSE)
dev.copy(png, file = "figure/plot_line_step_interval.png")
dev.off()
```

## Imputing missing values
### Imputing missing values
Note that there are a number of days/intervals where there are missing values 
(coded as NA). The presence of missing days may introduce bias into some
calculations or summaries of the data.
```{r}
# Calculate and report the total number of missing values in the dataset 
# (i.e. the total number of rows with NA
nas_total <- sum(is.na(df$steps))
```
The total number of **missing values** in the dataset is **`r nas_total`**.

Create a new dataset that is equal to the original dataset but with the missing
data filled in. The missing data are filled with the 
**mean for that 5-minute interval**.
```{r}
# Devise a strategy for filling in all of the missing values in the dataset. 
# The strategy does not need to be sophisticated. For example, you could use
# the mean/median for that day, or the mean for that 5-minute interval, etc.
# Merge data frame df & df_minute
df_replace <- merge(df, df_minute, by="interval")
# fill nas with mean by interval
df_replace <- df_replace %>%
       mutate(steps  = ifelse(is.na(steps), df_replace$mean, steps))
# sorted by date
df_replace <- df_replace[order(date),]
head(df_replace)
```
Make a histogram of the total number of steps taken each day and Calculate and 
report the mean and median total number of steps taken per day.
```{r hist_steps_day_nas_free}
## Make a histogram of the total number of steps taken each day and Calculate 
# and report the mean and median total number of steps taken per day without NA.
df_replace_day <- ddply(df_replace, .(date), summarize, 
                sum = sum(steps))

head(df_replace_day)
# Make a histogram of the total number of steps taken each day
hist(df_day$sum,
     main="Histogram of total number of steps taken per day without NAs", 
     xlab = "total number of steps taken each day",
     col = "#D6604D")

dev.copy(png, file = "figure/hist_steps_day_nas_free.png")
dev.off()

```
Calculate and report the mean and median total number of steps taken per day 
without NAs
```{r}
mean_step_day_ok <- mean(df_replace$steps)
median_step_day_ok <- median(df_replace$steps)
```
The **mean** of the total number of steps taken per day without NAs
is **`r mean_step_day_ok`**.
The **median** of the total number of steps taken per day without NAs
is **`r median_step_day_ok`**.

Do these values differ from the estimates from the first part of the assignment?
```{r}
dif_mean <- mean_step_day - mean_step_day_ok 
dif_median <- median_step_day - median_step_day_ok
```
The difference between the original mean and that of filling in Nas with the 
interval **mean** is **`r dif_mean`**.

The difference between the original mean and that of filling in Nas with the 
interval **median** is **`r dif_median`**.

What is the impact of imputing missing data on the estimates of the total daily 
number of steps?
```{r}
steps_original <- sum(df$steps, na.rm = TRUE)
steps_modified <- round(sum(df_replace$steps))
# remove scientific notation
options(scipen=999)
```
The total number of original daily steps is **`r steps_original `** and 
the total number of modified daily steps without NAs is **`r steps_modified `** 

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels – “weekday” and 
“weekend” indicating whether a given date is a weekday or weekend day.
```{r}
# is.weekend from library(chron)
df_replace <- df_replace %>%
        mutate(day_week  = ifelse(is.weekend(df_replace$date), 
                                  "weekend", "weekday"))
df_replace$day_week <- as.factor(df_replace$day_week)
str(df_replace)
```
Make a panel plot containing a time series plot (i.e. type = "l") 
of the 5-minute interval (x-axis) and the average number of steps taken, 
averaged across all weekday days or weekend days (y-axis).
```{r}
# summarize by day_week and interval
df_day_week <- ddply(df_replace, .(day_week, interval), summarize, 
                        mean_interval_day = mean(steps))
# xyplot from library(lattice)
with(df_day_week, xyplot(mean_interval_day ~ interval | day_week, 
                         type='l', 
                         layout = c(1, 2)))
dev.copy(png, file = "figure/lattice_plot_lines_mean_step_day_week.png")
dev.off()
```

