---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r setoptions,echo=TRUE}

```

## Loading and preprocessing the data
Load libraries and delete variables

```{r}
library(plyr)
library(dplyr)
library(data.table)
library(lubridate)
rm(list=ls())
```
Create data directory (if it does not exist previously) and unzip the .zip data
file in the "data" directory.
```{r}
# Create data folder if it does not exist
if (!file.exists("data")) {
        dir.create("data")
}
# Unzip file in "data" directory
unzip(zipfile = "activity.zip", exdir = "./data")
```

The data frame is created with fread and the data structure and type of 
each variable is checked.
```{r}
# Create a "Data Frame" with "fread"
df <- fread("./data/activity.csv")
head(df)
str(df)
summary(df)
```

## What is mean total number of steps taken per day?



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
